---
title: "A multinomial classification model for Volcano Eruptions"
author: "Gaurav Sharma"
date: '`r Sys.Date() `'
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE, dpi = 180, fig.width = 8, fig.height = 5)
library(tidyverse)
library(tidymodels)

```

### Lets laod the data

```{r}
volcano_raw <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv")
volcano_raw %>% 
    head(10) %>% 
    knitr::kable()
```

### Lets now explore the data
```{r}
volcano_raw %>% 
    count(primary_volcano_type, sort = T)
```

### It shows that there are 26 different types of valcanos. We will group them in three categories
```{r}
volcano_df <- volcano_raw %>% 
    transmute(volcano_type = case_when(str_detect(primary_volcano_type, "Stratovolcano") ~ "Stratovolcano",
                                       str_detect(primary_volcano_type, "Shield") ~ "Shield",
                                       TRUE ~ "Others"),
              volcano_number, latitude, longitude, tectonic_settings, elevation, major_rock_1) %>% 
    mutate_if(is.character, factor)
```

### Since we have spatial data, Lets see where these volcanos are situated on a map

```{r Volcano Map}
world <- map_data("world")

ggplot() +
    geom_map(data = world, map = world,
             aes(x = long, y = lat, map_id = region), color = "white", fill = "gray50", alpha = 0.2) +
    geom_point(data = volcano_df, aes(longitude, latitude, color = volcano_type), alpha = 0.5) +
    labs(x = NULL, y = NULL,
         title = "Different Type of Volcanic Eruptions around the world") +
    scale_x_continuous(labels = NULL) +
    scale_y_continuous(labels = NULL) 
    
```

### Rater than creating a split we will create resamples as we don't have much data
```{r}
volcano_boot <- volcano_df %>% 
    bootstraps()
volcano_boot
```

### We will create a recipe now. We will use smote analysis to overcome class imbalance
smote() function is a part of themis package. So, remember to load themis package before creating recipe.
```{r}
library(themis)

volcano_rec <- recipe(volcano_type~., data = volcano_df) %>% 
    update_role(volcano_number, new_role = "id") %>% 
    step_other(tectonic_settings) %>% 
    step_other(major_rock_1) %>% 
    step_dummy(tectonic_settings, major_rock_1) %>% 
    step_zv(all_predictors()) %>% 
    step_normalize(all_predictors()) %>% 
    step_smote(volcano_type)

volcano_prep <- prep(volcano_rec)

```

### Lets create a random forest model specification
```{r}
rf_spec <- rand_forest(trees = 1000) %>% 
    set_engine(engine = "ranger") %>% 
    set_mode(mode = "classification")
```

### Lets create a workflow
```{r}
volcano_wf <- workflow() %>% 
    add_recipe(recipe = volcano_rec) %>% 
    add_model(rf_spec)
```

### Lets train the model
```{r}
volcano_res <- fit_resamples(volcano_wf, 
              resamples = volcano_boot,
              control = control_resamples(save_pred = T, 
                                          verbose = T))
```

### Explore results
Let's see how the model 
```{r}
volcano_res %>% 
    collect_metrics()
```

```{r}
volcano_res %>% 
    collect_predictions() %>% 
    conf_mat(volcano_type, .pred_class)
```

```{r}
volcano_res %>% 
    collect_predictions() %>% 
    ppv(volcano_type, .pred_class)
```

```{r}
volcano_res %>% 
    collect_predictions() %>% 
    group_by(id) %>% 
    ppv(volcano_type, .pred_class)
```

```{r Vip plot}
library(vip)
rf_spec %>%
    set_engine("ranger", importance = "permutation") %>%
    fit(
        volcano_type ~ .,
        data = juice(volcano_prep) %>% select(-volcano_number) %>% janitor::clean_names()
    ) %>% 
    vip(geom = "point")
```


```{r}
volcano_pred <- volcano_res %>% 
    collect_predictions() %>% 
    mutate(correct = .pred_class == volcano_type) %>% 
    left_join(volcano_df %>% 
                  mutate(.row = row_number()))
```

# Lets see how we fared in terms of predictions by plotting our predictions again on a world map
```{r Volcano Pred Plot}
ggplot() +
    geom_map(data = world, map = world,
             aes(x = long, y = lat, map_id = region), color = "white", fill = "gray50", alpha = 0.5) +
    stat_summary_hex(data = volcano_pred, aes(longitude, latitude, z = as.integer(correct)), 
                     fun = "mean", alpha = 0.8, bins = 60) +
    labs(x = NULL, y = NULL,
         title = "On Average how accurate were our predictions across the world",
         fill = "Percent correct") +
    scale_x_continuous(labels = NULL) +
    scale_y_continuous(labels = NULL) +
    scale_fill_gradient(high = "cyan3", labels = scales::percent)
 
```






















